Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-13 paper/sections/proof-reconstruction.tex
[ paper ] working on section three.

Finally, we type-check the \Agda proof-term. If type-checking succeed, the \TSTP derivation delivered by \Metis was correct.
In that case \Agda outputs an interface file.
Otherwise, when type-checking fails, the failure must be investigated by the user, checking \Metis, \Athena, or even \Agda.

% We are responsible for errors committed in the translation from
% \TSTP to \Agda: bugs in the parsing module, lack of the coverage of
% \Metis' inference rules, and last but not least, the pretty print of
% the proof in \Agda code.

% In the rest of this section, we provide a formal description using
% type theory to build definitions and theorems for functions
% necessary to emulate \Metis' inference rules.
% Some conventions about notations. We will write the
% type judgments with ``$a : A$'' to denote that $a$ has type
% $\abbre{A}$. Types are written using small caps.
% For propositions, we will denote by $φ ≡ ψ$, the syntactic
% equivalence. We use the \Prop type for propositions and for theorems,
% the dependent type, $ Γ ⊢ φ$ where $Γ$ has type $\abbre{List} \Prop$ and
% $φ$ has type $\Prop$.

% % ...................................................................

% % section section_name (end)
% \subsection{Structural Recursion}
% \label{ssec:structural-recursion}

% To emulate \Metis' inference rules in type theory,
% we observed that transformations performed by some rules were given by \emph{general recursive} functions.
% But algorithms based on general recursions can not be translated
% directly in type theory since it is not a guarantee they terminate.

% Then, we solve the termination problem by modifying problematic functions
% to be \emph{structural recursive}~\cite{Coquand1992,Abel2002,Bove2005}.

% We use the same notion of \emph{structurally smaller} from
% \citeauthor{Abel2002}~\cite{Abel2002} to say, a recursive function is
% structurally recursive if it calls itself with only
% \emph{structurally smaller}
% arguments.

% To convert a general recursive call into a structurally recursive call,
% we use the \emph{Bounded Recursion} technique presented
% in~\cite{Bertot2004} but
% others found in \cite{Coquand1992,Abel2002,Bove2005} could be used.
% This technique converts a general recursive
% function into a structural recursive function by adding an additional
% argument for the function complexity, the bound for the number of times, the function needs to call itself to get the final result.

% Then, we add the natural number $n : \Nat$ in the input of the
% general recursive function, and perform the recursion over the structure of
% \Nat type. \abbre{Nat} type is the usual type for natural numbers with \zero and \suc as their constructors. We consider the common arithmetic operations ($+,\,-,\,*$) and numbers $0,1,2,\dots$ as representations of $\zero, \suc, \suc(\zero), \dots$.

% Then, the type-checker will
% recognize structurally smaller arguments by using the lexicographic
% order~\cite{Abel2002} for such kind of functions.

% In other words, we take a general recursive function $f: \Prop \to \Prop$.
% and we redefine that function as
% $f: \Prop \to \Nat \to \Prop$, and perform the recursion
% using induction on the structure of the \Nat type.

% Let us consider the following example to exhibit the technique
% mentioned above for an inner function in the \strip inference.

% \begin{align}
% \begin{split}
% \label{eq:unshunt-definition}
% &unshunt_{0}(φ) =\\
% &\hspace{3mm}\begin{cases}
% unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃),
%   & \text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
% unshunt_{0}(φ₁ ⇒ φ₂) ∧ unshunt_{0}(φ₁ ⇒ φ₃),
%   & \text{if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
% φ, &\text{otherwise.}
% \end{cases}
% \end{split}
% \end{align}

% From the definition above, the size of the argument, ((φ₁ ∧ φ₂) ⇒ φ₃),
% is not structurally smaller than the input formula, (φ₁ ⇒ (φ₂ ⇒ φ₃)), since ((φ₁ ∧ φ₂) ⇒ φ₃) is not a subformula of (φ₁ ⇒ (φ₂ ⇒ φ₃)).

% Thus, we define the function
% $unshunt_{1} : \Prop \to \Nat \to \Prop$
% by structural recursion in its second argument in
% Eq.~(\ref{eq:unshunt-structured}).

% \begin{align}
% \label{eq:unshunt-structured}
% \begin{split}
% &unshunt_{1}(φ, \zero)\hspace{3mm}= φ\\
% &unshunt_{1}(φ, \suc(n)) =\\
% &\hspace{3mm}\begin{cases}
% unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
%   &\text{if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
% unshunt_{1}(φ₁ ⇒ φ₂, n) ∧\ unshunt_{1}(φ₁ ⇒ φ₃, n),
%    &\text{if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
% \end{cases}
% \end{split}
% \end{align}

% As we mentioned above, the natural number in the second argument of the
% function used in Eq.~(\ref{eq:unshunt-structured}) stands for a complexity
% measure of the function, in this case $unshunt_{1}$ function.

% A \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
% and presented in \cite{VanDalen1994} as \emph{the rank for a
% proposition}, is a function $\Prop \to \Nat$ that assigns a natural
% number for a proposition φ as follows.

% \begin{equation*}
% \label{eq:rank-definition}
% rank(φ)=
% \begin{cases}
% 0, &\text{for atomic }φ \\
% rank(φ₁) + c_{¬},  &\text{if } φ ≡ \neg φ₁ \\
% \max{(rank(φ₁),rank(φ₂))} + c_{\square},
%   &\text{if } φ ≡ φ₁\,\square\, φ₂.
% \end{cases}
% \end{equation*}

% where $c_{¬}$ or $c_{\square}$ is a non-negative integer called the
% complexity of the connective.

% Recall, we want to maintain the same outputs as the original unshunt
% does in Eq.~(\ref{eq:unshunt-definition}) and the complexity measure of it,
% is a bound for the number of times this function can call itself.

% Therefore, we define an appropriate complexity measure
% $unshunt_{cm}: \Prop → \NN$ for our $unshunt_{1}$ function presented in
% Eq.~(\ref{eq:unshunt-rank-definition}).

% \begin{align}
% \begin{split}
% \label{eq:unshunt-rank-definition}
% &unshunt_{cm}(φ) =\\
% &\hspace{3mm}
% \begin{cases}
% unshunt_{cm}(φ₃) + 2, &\text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
% \max(unshunt_{cm}(φ₂),unshunt_{cm}(φ₃)) + 1,
%   &\text{if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
% 0, &\text{otherwise.}
% \end{cases}
% \end{split}
% \end{align}

% Finally, the curated version of $unshunt_{0}$ function is presented
% in Eq.~(\ref{eq:unshunt-fixed}).

% \begin{equation}
% \label{eq:unshunt-fixed}
% unshunt(\varphi) := unshunt_{1}(\varphi, unshunt_{cm}(\varphi)).
% \end{equation}

% For a more complete treatment of general recursion, and how to
% define well-founded and well-structural recursive functions, we refer
% the reader to Ref.~\cite{Bove2005}, where Bove and Capettra described
% not only general recursion, but they also treat mutual recursive
% function, among other kinds.

@@

% \subsection{Emulating \Metis' Inference Rules}
% \label{ssec:emulating-inferences}

% \subsubsection{Splitting a Goal.}
% \label{sssec:splitting-goal}

% The goal of a problem to be proved by \Metis is processed to generate new
% propositions, the new targets, \Metis calls these formulas
% \emph{subgoals}.

% The conjunction of these subgoals is a proposition equivalent to the
% original goal. Then, we find a refutation for each subgoal in the \Metis' \TSTP derivation.

% Nevertheless, \Metis does not provide neither an explanation for the user about the process in its \TSTP derivations nor in its documentation.

% In the following, we show how Theorem.~\ref{thm:thm-strip} proves the correctness of the process mentioned above. But first, let prove some lemmas for $unshunt$ function presented in Eq.~(\ref{eq:unshunt-fixed}),
% and $split$ function from the definition in Eq.~(\ref{eq:split-definition}).

% \begin{lemma}[lem-unshunt]
%   \label{lem:unshunt}
%   $Γ ⊢ unshunt(φ)$, then $Γ ⊢ φ$.
% \end{lemma}

% \begin{proof} Use induction on the cases defined by the outcome of the
% $unshunt$ function in Eq.~(\ref{eq:unshunt-structured}).
% When $n = 0$, by definition, $Γ ⊢ φ$.
% When $n = 1$, we apply the following theorem from~\cite{AgdaProp}

% \begin{equation*}
% \texttt{∧⇒-to-⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
% \end{equation*}

% For $n > 1$, we will use induction on the structure of second
% argument.

% (φ₁ ⇒ (φ₂ ⇒ φ₃))
% \begin{equation*}
%   \begin{bprooftree}
%   \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
%   \RightLabel{By def.}
%   \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
%   \RightLabel{lem-unshunt}
%   \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
%   \RightLabel{∧⇒-to-⇒⇒}
%   \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).}
%   \end{bprooftree}
% \end{equation*}

% Second, using the following theorem from \cite{AgdaProp},

% \begin{equation*}
%   \texttt{⇒∧-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).
% \end{equation*}

% ($φ₁ ⇒ (φ₂ ∧ φ₃)$)
% \begin{equation*}
% \scalebox{0.95}{
% ($\mathcal{D}_1$)
% \begin{bprooftree}
% \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n) ∧ unshunt(φ₁ ⇒ φ₃, n)$}
% \RightLabel{∧-proj₁}
% \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n)$}
% \RightLabel{By def.}
% \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n-1)$}
% \RightLabel{lem-unshunt}
% \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$,}
% \end{bprooftree}
% }
% \end{equation*}

% \begin{equation*}
% (\mathcal{D}_2)
% \begin{bprooftree}
% \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂) ∧ unshunt(φ₁ ⇒ φ₃)$}
% \RightLabel{∧-proj₂}
% \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n)$}
% \RightLabel{By def.}
% \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n-1)$}
% \RightLabel{lem-unshunt}
% \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$,}
% \end{bprooftree}
% \end{equation*}

% \begin{equation*}
%   \begin{bprooftree}
%   \AxiomC{$\mathcal{D}_1$}
%   \AxiomC{$\mathcal{D}_2$}
%   \RightLabel{∧-intro}
%   \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
%   \RightLabel{⇒∧⇒-to-⇒∧}
%   \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).}
%   \end{bprooftree}
% \end{equation*}
% \end{proof}

% Now, let us consider the $split$ function presented in
% Eq.~(\ref{eq:split-definition}) that yields a conjunction of subgoals from a proposition, the goal in \Metis.

% \begin{equation}
% \label{eq:split-definition}
% \mysplit(φ) =
% \begin{cases}
% unshunt(\mysplit(φ₁)) &\text{if }φ ≡ φ₁ ∧ φ₂\\
% ∧\ unshunt(φ₁ ⇒ \mysplit(φ₂)), &\\[2.5mm]
% unshunt(¬ φ₁ ⇒ \mysplit(φ₂)),
%   & \text{if }φ ≡ φ₁ ∨ φ₂\\[2.5mm]
% unshunt(φ₁ ⇒ \mysplit(φ₂)),
%   & \text{if }φ ≡ φ₁ ⇒ φ₂\\[2.5mm]
% unshunt(φ₁ ⇒ \mysplit(φ₂))
%   & \text{if }φ ≡ φ₁ ⇔ φ₂\\
% ∧\ unshunt(φ₂ ⇒ \mysplit(φ₁)),
%   &\\[2.5mm]
% unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
%   & \text{if }φ ≡ ¬ (φ₁ ∧ φ₂)\\[2.5mm]
% unshunt(\mysplit(¬ φ₁))
%   & \text{if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
% ∧\ unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂)),
%   &\\[2.5mm]
% unshunt(\mysplit(φ₁))
%   & \text{if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
% ∧\ unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
%   &\\[2.5mm]
% unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
%   & \text{if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
% ∧\ unshunt(¬ φ₂ ⇒ \mysplit(φ₁)),
%   &\\[2.5mm]
% unshunt(\mysplit(φ₁)),
%   & \text{if }φ ≡ ¬ ¬ φ₁\\
% ⊤,
%   & \text{if }φ ≡ ¬ ⊥\\
% ⊥,
%   & \text{if }φ ≡ ¬ ⊤.
% \end{cases}
% \end{equation}

% The definition above suffers termination problems since it is not
% a structural recursion. Therefore, we have reformulated Def.~\ref{eq:split-definition} to solve such a issue following the technique
% described in Section~\ref{ssec:structural-recursion}.
% Here and below, we refer to $split$ as the function after applying
% the process mentioned above.

% \begin{lemma}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
% \label{lem:lem-split}
% \end{lemma}

% \begin{proof} We now proceed by induction on the structure of the
% formula φ in the definition of Eq.~(\ref{eq:split-definition}).
% For a complete proof, we refer the reader to~\cite{AgdaMetis}.\\[3mm]

% ($φ ≡ φ₁ ∧ φ₂$). We get a proof for each conjunct and using the introduction rule for conjunction, the result follows.

% \begin{equation*}
% (\mathcal{D})\hspace{3mm}
%   \begin{bprooftree}
%   \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
%   \RightLabel{By def.}
%   \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
%   \RightLabel{∧-proj₁}
%   \UnaryInfC{$Γ ⊢ unshunt(φ₁)$}
%   \RightLabel{lem-unshunt}
%   \UnaryInfC{$Γ ⊢ φ₁$.}
%   \end{bprooftree}
% \end{equation*}

% \begin{equation*}
%   \begin{bprooftree}
%   \AxiomC{$\mathcal{D}$}
%   \UnaryInfC{Γ ⊢ φ₁}
%   \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
%   \RightLabel{By def.}
%   \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
%   \RightLabel{$∧$-proj$₂$}
%   \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
%   \RightLabel{lem-unshunt}
%   \UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
%   \RightLabel{$⇒$-elim}
%   \BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
%   \RightLabel{\thmsplit}
%   \UnaryInfC{$Γ ⊢ φ₂.$}
%   \end{bprooftree}
% \end{equation*}

% \vskip 3mm

>>>>

Finally, we type-check the \Agda proof-term. If type-checking
succeed, the \TSTP derivation delivered by \Metis was correct.
In that case \Agda outputs an interface file.
Otherwise, when type-checking fails, the failure must be
investigated by the user, checking for the error in \Metis, \Athena,
or in \Agda.

In the rest of this section, we provide a formal description using
type theory to build definitions and theorems for functions
necessary to emulate \Metis' inference rules. For this description,
we will write the type judgments with ``$a : A$'' to denote that $a$
has $\abbre{A}$ type. Types are written using small caps.
We use \Prop type for propositions and a the dependent type, $Γ ⊢ φ$ for theorems. In the latter type, $Γ$ has $\abbre{List}~\Prop$ type
and $φ$ has type $\Prop$. $\abbre{List}$ is the usual inductive type
for lists. Finally, we will denote by $φ ≡ ψ$, the syntactic
equivalence between propositions.

% ...................................................................

% section section_name (end)
\subsection{Structural Recursion}
\label{ssec:structural-recursion}

To emulate \Metis' inference rules in type theory,
we observed that transformations performed by some rules were given by \emph{general recursive} functions.
But algorithms based on general recursions can not be translated
directly in type theory since it is not a guarantee they terminate.
For that reason, we present in this section, a technique to avoid
such termination problems by modifying the recursive functions
to be \emph{structural recursive}.
% (see \cite{Coquand1992,Abel2002,Bove2005} for more details).

We adopt the notion of \emph{structurally smaller} from
\citeauthor{Abel2002}~\cite{Abel2002} to say, a recursive function is
structurally recursive if it calls itself with only
\emph{structurally smaller}
arguments.

General recursive calls can be translated into structurally recursive
calls by using for instance, the \emph{Bounded Recursion} technique
presented in~\cite{Bertot2004}. For a yet another method, we refer
the reader to \cite{Coquand1992,Abel2002,Bove2005}.
The former technique defines a new function based on the original
recursive function by adding an additional argument to the function.
The new argument is the \emph{bound}, a natural number given by the
function complexity. In other words, the added argument will store
the number of times the function needs to call itself to get the
expected outcome.

Henceforth, a natural number has \Nat type, and it is defined as
usual. That is, \zero and \suc are the data constructors and some
arithmetic operations are defined: ($+,\,-,\,*$). We include syntax
sugar for $\zero$, $\suc$, $\suc(\zero)$, $\cdots$, with the
classical representation: $1$, $2$, $\cdots$ as well.

Therefore, for a general recursive function $f: \Prop \to \Prop$
with termination problems, we states a new function $f^{*} : \Prop \to \Nat \to \Prop$ where all recursive calls must done by
using induction on the \Nat type structure.

\subsubsection{A practical Example.}
Let us consider the following example to exhibit the technique
mentioned above for an inner function present in the emulated version
of the \strip inference.

\begin{align}
\begin{split}
\label{eq:unshunt-definition}
&unshunt_{0}(φ) =\\
&\hspace{3mm}\begin{cases}
unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃),
  & \text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
unshunt_{0}(φ₁ ⇒ φ₂) ∧ unshunt_{0}(φ₁ ⇒ φ₃),
  & \text{if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
φ, &\text{otherwise.}
\end{cases}
\end{split}
\end{align}

From the definition above, the size of the argument, ((φ₁ ∧ φ₂) ⇒ φ₃),
is not structurally smaller than the input formula, (φ₁ ⇒ (φ₂ ⇒ φ₃)),
since ((φ₁ ∧ φ₂) ⇒ φ₃) is not a subformula of (φ₁ ⇒ (φ₂ ⇒ φ₃)).

Thus, we define the function
$unshunt_{1} : \Prop \to \Nat \to \Prop$
by structural recursion in its second argument in
Eq.~(\ref{eq:unshunt-structured}).

\begin{align}
\label{eq:unshunt-structured}
\begin{split}
&unshunt_{1}(φ, \zero)\hspace{3mm}= φ\\
&unshunt_{1}(φ, \suc(n)) =\\
&\hspace{3mm}\begin{cases}
unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
  &\text{if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\[2mm]
unshunt_{1}(φ₁ ⇒ φ₂, n)\\
∧\ unshunt_{1}(φ₁ ⇒ φ₃, n),
   &\text{if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{cases}
\end{split}
\end{align}

As we mentioned above, the natural number in the second argument of
the function used in Eq.~(\ref{eq:unshunt-structured}) stands for a
complexity measure of the function, in this case $unshunt_{1}$ function.

Recall, the \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
and presented in \cite{VanDalen1994} as \emph{the rank for a
proposition}, is a function $\Prop \to \Nat$ that assigns a natural
number for a proposition $φ$ like the following $rank$ function.
The symbols, $c_{¬}$ and $c_{\square}$ is a non-negative integer
number called the complexity of the connective.

\begin{equation*}
\label{eq:rank-definition}
rank(φ)=
\begin{cases}
0, &\text{for atomic }φ \\
rank(φ₁) + c_{¬},  &\text{if } φ ≡ \neg φ₁ \\
\max{(rank(φ₁),rank(φ₂))} + c_{\square},
  &\text{if } φ ≡ φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}

In the case of the $unshunt_0$ function in
Eq.~(\ref{eq:unshunt-definition}), we define the complexity function
$unshunt_{cm}: \Prop → \NN$ for our $unshunt_{1}$
presented in Eq.~(\ref{eq:unshunt-rank-definition}).

\begin{align}
\begin{split}
\label{eq:unshunt-rank-definition}
&unshunt_{cm}(φ) =\\
&\hspace{3mm}
\begin{cases}
unshunt_{cm}(φ₃) + 2, &\text{if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
\max(unshunt_{cm}(φ₂),unshunt_{cm}(φ₃)) + 1,
  &\text{if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
0, &\text{otherwise.}
\end{cases}
\end{split}
\end{align}

Finally, following the technique mentioned, we get a curated version of
$unshunt_{0}$ function presented in Eq.~(\ref{eq:unshunt-fixed}).

\begin{definition}[unshunt]
\begin{equation}
\label{eq:unshunt-fixed}
unshunt(\varphi) := unshunt_{1}(\varphi, unshunt_{cm}(\varphi)).
\end{equation}
\end{definition}

For a more complete treatment of general recursion, and how to
define well-founded and well-structural recursive functions, we refer
the reader to Ref.~\cite{Bove2005}, where Bove and Capettra described
not only general recursion, but they also treat mutual recursive
function, among other kinds.

@@

\subsection{Emulating \Metis' Inference Rules}
\label{ssec:emulating-inferences}

\subsubsection{Splitting a Goal.}
\label{sssec:splitting-goal}

 To prove a goal, \Metis splits the goal into
disjoint cases. This process produces a list of new subgoals, the
conjunction of these subgoals implies the goal. Then, a proof of the
goal becomes in smaller proofs, one refutation for each subgoal.
These subgoals are introduced in the \TSTP derivation with the \strip
inference rule.

\begin{verbatim}
fof(goal, conjecture, p & r & q).
fof(subgoal_0, plain, p, inference(strip, [], [goal])).
fof(subgoal_1, plain, p => r, inference(strip, [], [goal])).
fof(subgoal_2, plain, (p & r) => q, inference(strip, [], [goal])).
\end{verbatim}

In the following, we show how Theorem~\ref{thm:thm-strip} proves the
correctness of the process mentioned above. But first, let us prove some
lemmas for $unshunt$ function presented in Eq.~(\ref{eq:unshunt-fixed}),
and $split$ function from the definition in Eq.~(\ref{eq:split-definition}).

\begin{lemma}[lem-unshunt]
  \label{lem:unshunt}
  $Γ ⊢ unshunt(φ)$, then $Γ ⊢ φ$.
\end{lemma}

\begin{proof} Use induction on the cases defined by the outcome of the
$unshunt$ function in Eq.~(\ref{eq:unshunt-structured}).
\begin{itemize}
	\item When $n = 0$, by definition, $Γ ⊢ φ$.
	\item When $n = 1$, we apply the following theorem from~\cite{AgdaProp}
		\begin{equation*}
		\texttt{∧⇒-to-⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
		\end{equation*}
  \item For $n > 1$, we will use induction on the structure of the second
			  argument.

\begin{itemize}
\item $(φ₁ ⇒ (φ₂ ⇒ φ₃))$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒-to-⇒⇒}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).}
  \end{bprooftree}
\end{equation*}

\item $(φ₁ ⇒ (φ₂ ∧ φ₃))$
\begin{equation*}
\scalebox{0.95}{
($\mathcal{D}_1$)
\begin{bprooftree}
\AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n) ∧ unshunt(φ₁ ⇒ φ₃, n)$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n-1)$}
\RightLabel{lem-unshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$,}
\end{bprooftree}
}
\end{equation*}

\begin{equation*}
(\mathcal{D}_2)
\begin{bprooftree}
\AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂) ∧ unshunt(φ₁ ⇒ φ₃)$}
\RightLabel{∧-proj₂}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n-1)$}
\RightLabel{lem-unshunt,}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
\end{bprooftree}
\end{equation*}
Finally, using the theorem, \texttt{⇒∧⇒-to-⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒-to-⇒∧.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}

\end{itemize}
\end{itemize}

\end{proof}

Let us now consider the $split$ function presented in
Eq.~(\ref{eq:split-definition}) that yields the conjunction of subgoals that implies the goal in a \Metis \TSTP derivation.

\begin{definition}[split]
\begin{equation}
\label{eq:split-definition}
\mysplit(φ) =
\begin{cases}
unshunt(\mysplit(φ₁)) &\text{if }φ ≡ φ₁ ∧ φ₂\\
∧\ unshunt(φ₁ ⇒ \mysplit(φ₂)), &\\[2.5mm]
unshunt(¬ φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ∨ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂)),
  & \text{if }φ ≡ φ₁ ⇒ φ₂\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(φ₂))
  & \text{if }φ ≡ φ₁ ⇔ φ₂\\
∧\ unshunt(φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  & \text{if }φ ≡ ¬ (φ₁ ∧ φ₂)\\[2.5mm]
unshunt(\mysplit(¬ φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
∧\ unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁))
  & \text{if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
∧\ unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
  &\\[2.5mm]
unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
  & \text{if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
∧\ unshunt(¬ φ₂ ⇒ \mysplit(φ₁)),
  &\\[2.5mm]
unshunt(\mysplit(φ₁)),
  & \text{if }φ ≡ ¬ ¬ φ₁\\
⊤,
  & \text{if }φ ≡ ¬ ⊥\\
⊥,
  & \text{if }φ ≡ ¬ ⊤.
\end{cases}
\end{equation}
\end{definition}

The definition above suffers termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation above to solve such a issue following the technique
described in Section~\ref{ssec:structural-recursion}.

Here and below, we refer to $split$ as the function after applying
the process mentioned above.

\begin{lemma}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
\label{lem:lem-split}
\end{lemma}

\begin{proof} We now proceed by induction on the structure of the
formula φ in the definition of Eq.~(\ref{eq:split-definition}).
For a complete proof, we refer the reader to~\cite{AgdaMetis}.\\[3mm]

($φ ≡ φ₁ ∧ φ₂$). We get a proof for each conjunct and using the introduction rule for conjunction, the result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁)$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁$.}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ ⊢ φ₂.$}
  \end{bprooftree}
\end{equation*}

\vskip 3mm

($φ ≡ φ₁ ⇒ φ₂$).
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
  \RightLabel{By def.}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$.}
  \end{bprooftree}
\end{equation*}
\vskip 3mm

(φ ≡ ¬ (φ₁ ∨ φ₂)). We show $Γ ⊢ ¬ φ₁$ and $Γ ⊢ ¬ φ₂$. Using their conjunction by the introduction rule, and De Morgan Law, the results follows.
